---
title: "project 2"
author: "CanPeng"
date: "3/18/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(mlbench)
library(tidyverse)
```


### Data preparation 
#Load Data

```{r}
# load the data set
data(BreastCancer)
# some algorithms don't like missing values, so remove rows with missing values
BreastCancer <- na.omit(BreastCancer) 
# remove the unique identifier, which is useless and would confuse the machine learning algorithms
BreastCancer$Id <- NULL 
# check the class of each varibles
str(BreastCancer)
```

#partition the data

```{r}
# partition the data set for 80% training and 20% evaluation
set.seed(2)
ind <- sample(2, nrow(BreastCancer), replace = TRUE, prob=c(0.8, 0.2))
train <- BreastCancer[ind==1,]
valid <- BreastCancer[ind==2,]
```

### Create multiple models using different classifiers/algorithms 

#1 Decision trees
```{r}
library(rpart)
library(rpart.plot)
x.rp <- rpart(Class ~ ., data=train)
#plot(x.rp, main="Decision tree created using rpart") 
prp(x.rp, type = 1, extra = 1, split.font = 1, varlen = -10)  
```
```{r}
#prediction
# predict classes for the evaluation data set
x.rp.pred <- predict(x.rp, type="class", newdata=valid)
# score the evaluation data set (extract the probabilities)
x.rp.prob <- predict(x.rp, type="prob", newdata=valid)
table(x.rp.pred,valid$Class)
```
# Leave-1-Out Cross Validation (LOOCV)
```{r}
ans <- numeric(nrow(BreastCancer))
for (i in 1:nrow(BreastCancer)) {
  mytree <- rpart(Class ~ ., BreastCancer[-i,])
  mytree.pred <- predict(mytree,BreastCancer[i,],type="class")
  ans[i] <- mytree.pred
}
ans <- factor(ans,labels=levels(BreastCancer$Class))
table(ans,BreastCancer$Class)
```

#2 condition inference trees
#  create model using conditional inference trees
```{r}
require(party)
x.ct <- ctree(Class ~ ., data=train)
plot(x.ct, main="Decision tree created using condition inference trees") 

x.ct.pred <- predict(x.ct, newdata=valid) #ensemble
x.ct.prob <-  1- unlist(treeresponse(x.ct, valid), use.names=F)[seq(1,nrow(valid)*2,2)]
table(x.ct.pred,valid$Class)
```

#3 Random Forests

```{r}
x.cf <- cforest(Class ~ ., data=train, control = cforest_unbiased(mtry = 9)) 

x.cf.pred <- predict(x.cf, newdata=valid)
x.cf.prob <-  1- unlist(treeresponse(x.cf, valid), use.names=F)[seq(1,nrow(valid)*2,2)]
table(x.cf.pred,valid$Class)
```

#4 bagging
# create model using bagging (bootstrap aggregating)
```{r}
require(ipred)
x.ip <- bagging(Class ~ ., data=train) 
x.ip.pred <- predict(x.ip, newdata=valid)
x.ip.prob <- predict(x.ip, type="prob", newdata=valid)
table(x.ip.pred,valid$Class)
```

#5.SVM
# create model using svm (support vector machine)
```{r}
require(e1071)
# svm requires tuning
x.svm.tune <- tune(svm, Class~., data = train,
                   ranges = list(gamma = 2^(-8:1), cost = 2^(0:4)),
                   tunecontrol = tune.control(sampling = "fix")) 
# display the tuning results (in text format)
x.svm.tune #note the gamma and cost
# If the tuning results are on the margin of the parameters (e.g., gamma = 2^-8), 
# then widen the parameters.
# I manually copied the cost and gamma from console messages above to parameters below.
x.svm <- svm(Class~., data = train, cost=1, gamma=0.03125, probability = TRUE) 
x.svm.pred <- predict(x.svm, type="class", newdata=valid)
x.svm.prob <- predict(x.svm, type="prob", newdata=valid, probability = TRUE)
table(x.svm.pred,valid$Class)
```

#7. neural network
```{r}
library(nnet)
x.nn <- nnet(Class~., data = train,size=1) 
x.nn.pred <- predict(x.nn,valid,type="class")
x.nn.prob <- predict(x.nn,valid,type="raw")
table(x.nn.pred,valid$Class)
```

#8 QDA
#Quadratic Discriminant Analysis
```{r}
library(MASS)
library(dplyr)
train.num <- train %>% dplyr::select(-Class) %>% mutate_if(is.factor,as.character)%>% mutate_if(is.character,as.numeric) #change the class
train.num$Class <- train$Class
valid.num <- valid%>%dplyr::select(-Class) %>% mutate_if(is.factor,as.character)%>% mutate_if(is.character,as.numeric) #change the class
valid.num$Class <- valid$Class
str(train.num)

x.qda <- qda(Class~., data = train.num) 
x.qda.pred <- predict(x.qda, valid.num)$class
table(x.qda.pred,valid.num$Class)
```

#9 RDA
#Regularised Discriminant Analysis
```{r}
library(klaR)
x.rda <- rda(Class~., data = train)
x.rda.pred <- predict(x.rda, valid)$class
table(x.rda.pred,valid$Class)
```

## ROC
## plot ROC curves to compare the performance of the individual classifiers

# Output the plot to a PNG file for display on web.  To draw to the screen, 
# comment this line out.
#png(filename="roc_curve_5_models.png", width=700, height=700)


```{r}
#load the ROCR package which draws the ROC curves
require(ROCR)

# rptree
# create an ROCR prediction object from rpart() probabilities
x.rp.prob.rocr <- prediction(x.rp.prob[,2], BreastCancer[ind == 2,'Class'])
# prepare an ROCR performance object for ROC curve (tpr=true positive rate, fpr=false positive rate)
x.rp.perf <- performance(x.rp.prob.rocr, "tpr","fpr")

# ctree
x.ct.prob.rocr <- prediction(x.ct.prob, BreastCancer[ind == 2,'Class'])
x.ct.perf <- performance(x.ct.prob.rocr, "tpr","fpr")

# cforest
x.cf.prob.rocr <- prediction(x.cf.prob, BreastCancer[ind == 2,'Class'])
x.cf.perf <- performance(x.cf.prob.rocr, "tpr","fpr")

# bagging
x.ip.prob.rocr <- prediction(x.ip.prob[,2], BreastCancer[ind == 2,'Class'])
x.ip.perf <- performance(x.ip.prob.rocr, "tpr","fpr")

# svm
x.svm.prob.rocr <- prediction(attr(x.svm.prob, "probabilities")[,2], BreastCancer[ind == 2,'Class'])
x.svm.perf <- performance(x.svm.prob.rocr, "tpr","fpr")

# nn
x.nn.prob.rocr <- prediction(x.nn.prob, BreastCancer[ind == 2,'Class'])
x.nn.perf <- performance(x.nn.prob.rocr, "tpr","fpr")

```


```{r}
# plot it
plot(x.rp.perf, col=2, main="ROC curves comparing classification performance of six machine learning models") 
# Draw a legend
legend(0.6, 0.6, c('rpart', 'ctree', 'cforest','bagging','svm', 'neural network'),2:6)
plot(x.ct.perf, col=3, add=TRUE)
plot(x.cf.perf, col=4, add=TRUE)
plot(x.ip.perf, col=5, add=TRUE)
plot(x.svm.perf, col=6, add=TRUE)
plot(x.nn.perf, col=7, add=TRUE)
```

### Ensemble: creating an ensemble for combining all classifiers

```{r}
classifier <- data.frame(cbind(x.rp.pred, x.ct.pred, x.cf.pred, x.ip.pred, x.svm.pred ,x.nn.pred))

names(classifier) <-c('recursive.tree','conditional.inference.tree','random.forest','bootstrap','svm','neutral.network')
levels(classifier$neutral.network) =c('1','2')

classifier <-classifier%>% sapply(FUN = function(x)(ifelse(x=='1',0,1)))
classifier<- addmargins(classifier, margin = 2) # table/arragy, margin =2 aggregate by col 
classifier <- data.frame(classifier)
classifier$predition <- ifelse(classifier$Sum >=5, 'malignant','benign')
head(classifier)
```

```{r}
library(ggplot2)
table(classifier$predition, valid$Class)
```

#confusion matrix
```{r}
library(caret)
confusionMatrix(as.factor(classifier$predition), valid$Class, positive = 'malignant')
```